---
title: "Exposure income distributions"
author: "Jeanette Choong"
date: "2023-11-21"
output: html_document
---
```{r setup, include=FALSE}
library(raster)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(scales)
library(data.table)
```

## ABOUT

This document seeks to visualise how income levels influence flood exposure in the Philippines. 

We look at this using a 1-in-100 year coastal flood hazard map for 2050 (with sea level rise) and 2020 (baseline scenario).

## 1. SET UP

### 1.1 LOAD DATA

We start with the poverty line. According to the Philippine Statistical Authority, the poverty threshold for a household of five was 12,030 PHP/month in 2021.
The international poverty line is US$2.15/day in 2017 PPP. Below are the conversions of the international and national poverty lines in thousand pesos per year for a household of five. 

```{r poverty line}
pov_line_natl <- 12030 * 12 /1000 # for a hh in 000 PHP/year 
pov_line_intl <- 47.8 * 5 * 365 /1000 # for a hh in 000 PHP/year 

paste(pov_line_natl, "|", pov_line_intl)
```

```{r rasters}
flood_rast20 <- raster("/Users/jeancjw/Documents/00_Data/Coastal flood maps/Philippines/RCP45/Y2020/merged/merge_Y2020_RP100.tif") # for year 2020
flood_rast <- raster("/Users/jeancjw/Documents/00_Data/Coastal flood maps/Philippines/RCP45/Y2050/merged/merge_Y2050_RP100.tif") # year 2050
labour_income_rast <- raster("/Users/jeancjw/Documents/00_Data/Income expenditure savings 2021/Labour income rasters/labour_income_merged.tif")
building_rast <- raster("/Users/jeancjw/Documents/00_Data/Buildings/buildingcount_90m.tif")
```

### 1.2 CONVERT RASTERS TO DATAFRAME

```{r}
# Get pixels where there are buildings
flood_rast[is.na(building_rast)] <- NA
labour_income_rast[is.na(building_rast)] <- NA


# Convert rasters to points 
flood_pts <- rasterToPoints(flood_rast)
income_pts <- rasterToPoints(labour_income_rast)
building_pts <- rasterToPoints(building_rast)


# Check if 'x' and 'y' coordinates are the same for every row
coordinates_match <- all.equal(flood_pts[, c("x", "y")], income_pts[, c("x", "y")]) # check building_pts as well. This code cannot compare 3 data frames.

if (is.logical(coordinates_match) && all(coordinates_match)) {
  print("Coordinates are the same for every row.")
} else {
  stop("Coordinates are not the same for every row. Check your data.")
}
head(income_pts)
```

Since the coordinates for each row are the same in all three data frames, we can simply bind the three together without using merge(), because we don't really need the x y coords for this analysis. 

```{r}
data <- cbind(flood_pts[, c("x", "y", "layer")], income_pts[, "layer"], building_pts[, "layer"]) #'layer' is just the name of the col containing flood depth, income and building values in each df/matrix

head(data) # check
```
Now we join building count data

```{r}
# Rename columns for clarity
 colnames(data) <- c("x", "y", "depth", "income", "buildings")

head(data) # check
```

Now we want to expand the data frame such that each row represents one building.

```{r}
data <- as.data.table(data) # convert matrix to df
data <- data %>% 
  select(-x,-y) # remove x y coordinates; we don't need this

# Replicate rows c(income, depth) based on number of buildings 
## note: requires data to be in data.table format; doesn't work with data frame
data <- data[, .(income = rep(income, buildings), 
                depth = rep(depth, buildings)), by = 1:nrow(data)] 

data <- data %>% 
  filter(depth <= 435.7778) %>% # some weird bug in R that's causing the max flood depth to appear as 20k even though the max is only 435.78cm
  select(-nrow) # Don't need this column

data # check data
```

## 2. PLOT PDF

### 2.1 Y2050

We start by categorising each building into poor and non-poor, based on which side of the poverty line they fall under.

In the first case, we include even buildings that are unaffected.

#### 2.1.1 NATIONAL POVERTY LINE

```{r national pov line}
data %>% 
  mutate(income_category = ifelse(income <= pov_line_natl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 14, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```
Since there are too many unaffected buildings, it's difficult to see any pattern amongst the affected buildings, so what we want to do next is to exclude buildings where flood depth = 0. 

However, it's worth noting that fewer poor households are not exposed to floods compared to non-poor households. This could also be due to the way we have categorised poor and non-poor, meaning the bulk of the households fall under non-poor. 

```{r national pov line - exposed only}
data %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_natl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 4, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```

This figure shows the probability distribution of flood exposure of households above and below the poverty line in the Philippines, for a 1-in-100 year coastal flood in 2050. 
We can see here that generally, households living under the poverty line are more exposed at higher flood depths, which those living above the poverty line tend to face lower flood exposures


#### 2.1.2 INTERNATIONAL POVERTY LINE

```{r international pov line}
data %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 14, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```

```{r international pov line - exposed only}
data %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 4, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```


#### 2.1.3 INCOME QUANTILES

The problem with categorising by poverty thresholds is that there tends to be only a very small population who fall under these thresholds. This means even the poor who aren't living under the poverty line are classified as non-poor, which may undermine the real exposure on this "in between" group. 

```{r quantiles top bottom}
data %>%
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Bottom 20%",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Top 20%"))))) %>%
   mutate(income_category = factor(income_category, levels = c("Top 20%", "Q4", "Q3", "Q2", "Bottom 20%"))) %>% # reorder so that bottom 20% appears as the layer in front
   filter(income_category %in% c("Bottom 20%", "Top 20%")) %>% # plot only lowest and highest incomes
  ggplot(aes(x = depth, fill = income_category)) +
  stat_density(alpha = 0.7, 
               adjust = 14, 
               position = "identity") +
  scale_fill_manual(values = c("Bottom 20%" = "#294475", "Top 20%" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income category"
  )
```



```{r quantiles top bottom - exposed only}
data %>%
  filter(depth > 0) %>%
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Bottom 20%",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Top 20%"))))) %>%
  mutate(income_category = factor(income_category, levels = c("Top 20%", "Q4", "Q3", "Q2", "Bottom 20%"))) %>% # reorder so that bottom 20% appears as the layer in front
   filter(income_category %in% c("Bottom 20%", "Top 20%")) %>% # plot only lowest and highest incomes
  ggplot(aes(x = depth, fill = income_category)) +
  stat_density(alpha = 0.7, adjust = 4, position= "identity") +
  scale_fill_manual(values = c("Bottom 20%" = "#294475", "Top 20%" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income category"
  )
```

Now we plot the same thing for all quantiles

```{r all quantiles}
data %>%
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Q1",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Q5"))))) %>%
  ggplot(aes(x = depth)) +
  stat_density(alpha = 0.5, 
               adjust = 14, 
               position = "identity") +
  # scale_fill_manual(values = c("Q1" = "#294475", "Q2" = "#F0E442", "Q3" = "#56B4E9", "Q4" = "#CC79A7", "Q5" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income quantile"
  ) + 
  facet_wrap(~income_category, scales="fixed")
```

```{r}
data %>%
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Q1",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Q5"))))) %>%
  ggplot(aes(x = depth)) +
  stat_density(alpha = 0.5, 
               adjust = 4, 
               position = "identity") +
  # scale_fill_manual(values = c("Q1" = "#294475", "Q2" = "#F0E442", "Q3" = "#56B4E9", "Q4" = "#CC79A7", "Q5" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income quantile"
  ) + 
  facet_wrap(~income_category, scales="fixed")
```
This plot is interesting because Q2 - Q5 seem to have similar exposure levels, but Q1 shows a disproportionately high exposure to high flood levels. 

### 2.2 Y2020

```{r}
# Get pixels where there are buildings
flood_rast20[is.na(building_rast)] <- NA

# Convert rasters to points 
flood_pts20 <- rasterToPoints(flood_rast20)


# Check if 'x' and 'y' coordinates are the same for every row
coordinates_match <- all.equal(flood_pts20[, c("x", "y")], income_pts[, c("x", "y")]) # check building_pts as well. This code cannot compare 3 data frames.

if (is.logical(coordinates_match) && all(coordinates_match)) {
  print("Coordinates are the same for every row.")
} else {
  stop("Coordinates are not the same for every row. Check your data.")
}
```
Since the coordinates for each row are the same in all three data frames, we can simply bind the three together without using merge(), because we don't really need the x y coords for this analysis. 

```{r}
data20 <- cbind(flood_pts20[, c("x", "y", "layer")], income_pts[, "layer"], building_pts[, "layer"]) #'layer' is just the name of the col containing flood depth, income and building values in each df/matrix

head(data20) # check
```
Now we join building count data

```{r}
# Rename columns for clarity
 colnames(data20) <- c("x", "y", "depth", "income", "buildings")

head(data20) # check
```

Now we want to expand the data frame such that each row represents one building.

```{r}
data20 <- as.data.table(data20) # convert matrix to df
data20 <- data20 %>% 
  select(-x,-y) # remove x y coordinates; we don't need this

# Replicate rows c(income, depth) based on number of buildings 
## note: requires data20 to be in data.table format; doesn't work with data20 frame
data20 <- data20[, .(income = rep(income, buildings), 
                depth = rep(depth, buildings)), by = 1:nrow(data20)] 

data20 <- data20 %>% 
  filter(depth <= 435.7778) %>% # some weird bug in R that's causing the max flood depth to appear as 20k even though the max is only 435.78cm
  select(-nrow) # Don't need this column

data20 # check data20
```


#### 2.2.1 NATIONAL POVERTY LINE

```{r national pov line}
data20 %>% 
  mutate(income_category = ifelse(income <= pov_line_natl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 14, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```
Since there are too many unaffected buildings, it's difficult to see any pattern amongst the affected buildings, so what we want to do next is to exclude buildings where flood depth = 0. 

However, it's worth noting that fewer poor households are not exposed to floods compared to non-poor households. This could also be due to the way we have categorised poor and non-poor, meaning the bulk of the households fall under non-poor. 

```{r national pov line - exposed only}
data20 %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_natl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 4, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```

This figure shows the probability distribution of flood exposure of households above and below the poverty line in the Philippines, for a 1-in-100 year coastal flood in 2050. 
We can see here that generally, households living under the poverty line are more exposed at higher flood depths, which those living above the poverty line tend to face lower flood exposures

#### 2.2.2 INTERNATIONAL POVERTY LINE 

```{r international pov line}
data20 %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 14, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```

```{r international pov line - exposed only}
data20 %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
ggplot(aes(x = depth, fill = income_category))+
  stat_density(alpha = 0.7, adjust = 4, position = "identity") + 
  scale_fill_manual(values = c("Below poverty line" = "#294475", "Above poverty line" = "#cf5519")) + 
  theme_bw()+
  labs(x="Flood depth (cm)",
       y= "Frequency", 
       fill = "Income category") 
```


#### 2.2.3 INCOME QUANTILES

The problem with categorising by poverty thresholds is that there tends to be only a very small population who fall under these thresholds. This means even the poor who aren't living under the poverty line are classified as non-poor, which may undermine the real exposure on this "in between" group. 

```{r quantiles top bottom}
data20 %>%
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Bottom 20%",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Top 20%"))))) %>%
   mutate(income_category = factor(income_category, levels = c("Top 20%", "Q4", "Q3", "Q2", "Bottom 20%"))) %>% # reorder so that bottom 20% appears as the layer in front
   filter(income_category %in% c("Bottom 20%", "Top 20%")) %>% # plot only lowest and highest incomes
  ggplot(aes(x = depth, fill = income_category)) +
  stat_density(alpha = 0.7, 
               adjust = 14, 
               position = "identity") +
  scale_fill_manual(values = c("Bottom 20%" = "#294475", "Top 20%" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income category"
  )
```



```{r quantiles top bottom - exposed only}
data20 %>%
  filter(depth > 0) %>%
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Bottom 20%",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Top 20%"))))) %>%
  mutate(income_category = factor(income_category, levels = c("Top 20%", "Q4", "Q3", "Q2", "Bottom 20%"))) %>% # reorder so that bottom 20% appears as the layer in front
   filter(income_category %in% c("Bottom 20%", "Top 20%")) %>% # plot only lowest and highest incomes
  ggplot(aes(x = depth, fill = income_category)) +
  stat_density(alpha = 0.7, adjust = 4, position= "identity") +
  scale_fill_manual(values = c("Bottom 20%" = "#294475", "Top 20%" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income category"
  )
```

Now we plot the same thing for all quantiles

```{r all quantiles}
data20 %>%
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Q1",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Q5"))))) %>%
  ggplot(aes(x = depth)) +
  stat_density(alpha = 0.5, 
               adjust = 14, 
               position = "identity") +
  # scale_fill_manual(values = c("Q1" = "#294475", "Q2" = "#F0E442", "Q3" = "#56B4E9", "Q4" = "#CC79A7", "Q5" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income quantile"
  ) + 
  facet_wrap(~income_category, scales="fixed")
```

```{r}
data20 %>%
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Q1",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Q5"))))) %>%
  ggplot(aes(x = depth)) +
  stat_density(alpha = 0.5, 
               adjust = 4, 
               position = "identity") +
  # scale_fill_manual(values = c("Q1" = "#294475", "Q2" = "#F0E442", "Q3" = "#56B4E9", "Q4" = "#CC79A7", "Q5" = "#cf5519")) +
  theme_bw() +
  labs(
    x = "Flood depth (cm)",
    y = "Frequency",
    fill = "Income quantile"
  ) + 
  facet_wrap(~income_category, scales="fixed")
```

## 3. PLOT CDF

### 3.1 Y2050 

```{r cdf 2050 - natl}
data %>%  
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_natl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=depth, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Flood depth (cm)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw()
```

In this plot the curve for households at or below the poverty threshold is consistently below that of the curve for households above the poverty line. This suggests that across all observed flood depths, those living at or under the poverty threshold are generally more exposed to coastal floods. The gap between the two curves highlights this inequity in exposure. 

```{r cdf 2050 - intl}
data %>%  
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=depth, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Flood depth (cm)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw()
```

The vertical distance between both curves show the difference in cumulative frequency at any given flood depth.
The larger the gap, the greater the inequality in exposure to that flood depth. 

We therefore want to compute the maximum distance between the two curves. 
#### 3.1.1 Max vertical dist

```{r}
low_income <- data %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Below poverty line") %>% 
  arrange(depth)
high_income <- data %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Above poverty line") %>% 
  arrange(depth)


cdf_low <- ecdf(low_income$depth)
cdf_high <- ecdf(high_income$depth)

data %>%  
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=depth, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Flood depth (cm)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw() + 
  geom_vline(xintercept = 124.321, color = "grey", size = 0.5)
```

```{r}
depth_values <- seq(min(data$depth), max(data$depth), length.out = 1000)
differences <- cdf_high(depth_values) - cdf_low(depth_values)
abs_difference <- abs(differences)
l1norm50 <- sum(abs_difference)
cat("L1 Norm:", l1norm50, "\n")


max_difference <- max(differences)
max_difference_depth <- depth_values[which.max(differences)]


cat("Maximum Difference:", max_difference, "\n")
cat("Corresponding Flood Depth:", max_difference_depth, "\n")
```





### 3.2 Y2020 

```{r}
low_income20 <- data20 %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Below poverty line") %>% 
  arrange(depth)
high_income20 <- data20 %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Above poverty line") %>% 
  arrange(depth)


cdf_low <- ecdf(low_income20$depth)
cdf_high <- ecdf(high_income20$depth)

data20 %>%  
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=depth, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Flood depth (cm)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw() 
```

```{r}
depth_values <- seq(min(data20$depth), max(data20$depth), length.out = 1000)
differences <- cdf_high(depth_values) - cdf_low(depth_values)
l1norm20 <- sum(abs(differences))

l1norm20
max_difference <- max(differences)
max_difference_depth <- depth_values[which.max(differences)]

cat("Maximum Difference:", max_difference, "\n")
cat("Corresponding Flood Depth:", max_difference_depth, "\n")
```

```{r}
data <- data %>%
  mutate(year = rep("2050", nrow(.))) %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Bottom 20%",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Top 20%"))))) %>%
  mutate(income_category = factor(income_category, levels = c("Top 20%", "Q4", "Q3", "Q2", "Bottom 20%")))


data20 <- data20%>%
  mutate(year = rep("2020", nrow(.))) %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= quantile(income, 0.2, na.rm = TRUE), "Bottom 20%",
                                  ifelse(income <= quantile(income, 0.4, na.rm = TRUE), "Q2",
                                         ifelse(income <= quantile(income, 0.6, na.rm = TRUE), "Q3",
                                                ifelse(income <= quantile(income, 0.8, na.rm = TRUE), "Q4", "Top 20%"))))) %>%
  mutate(income_category = factor(income_category, levels = c("Top 20%", "Q4", "Q3", "Q2", "Bottom 20%")))

data_all <- rbind(data, data20)

data_all <- bind_rows(
  mutate(data_all %>% filter(year == 2020),
         year = factor(year),
         color = "#cf5519"),  # Color for 2020
  mutate(data_all %>% filter(year == 2050),
         year = factor(year),
         color = "#294475")  # Color for 2050
)

data_all %>%
  filter(depth > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  ggplot(aes(x = depth, linetype = income_category, color = year)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Flood depth (cm)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Year") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()


```

Calculate maximum distance for this curve (), then send to David

Do something similar to welfare loss to look at the difference between


```{r without inequity}
data_all %>%
  filter(depth > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>%
  filter(year == "2050") %>%
  ggplot(aes(x = depth, color = income_category)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Flood depth (cm)",
       y = "Cumulative frequency") +
  # scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("white", "black")) +
  theme_bw()
 
```


### 3.3 COMBINE BOTH YEARS


```{r}
data20_cat <- data20 %>% 
  filter(depth > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line"))

data50_cat <- data %>% 
  filter(depth > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line"))

# separate categories
data20_below <- data20_cat %>%  
  filter(income_category=="Below poverty line") %>% 
  arrange(depth)
data50_below <- data50_cat %>%  
  filter(income_category=="Below poverty line") %>% 
  arrange(depth)

data20_above<- data20_cat %>%  
  filter(income_category=="Above poverty line") %>% 
  arrange(depth)
data50_above <- data50_cat %>%  
  filter(income_category=="Above poverty line") %>% 
  arrange(depth)


# Now calculate max diff between the two years for each income category
cdf_low_20 <- ecdf(data20_below$depth)
cdf_low_50 <- ecdf(data50_below$depth)

cdf_high_20 <- ecdf(data20_above$depth)
cdf_high_50 <- ecdf(data50_above$depth)

## Calculate max difference 
depth_values <- seq(min(data20_below$depth), max(data20_below$depth), length.out = 1000)
differences <- cdf_low_20(depth_values) - cdf_low_50(depth_values)


max_difference <- max(differences)
max_difference_depth <- depth_values[which.max(differences)]

cat("Maximum Difference:", max_difference, "\n")
cat("Corresponding Flood Depth:", max_difference_depth, "\n")
```

## 4. PLOT ASSET LOSSES

```{r}
# LOAD DATA
AL_rast_20 <- raster("/Users/jeancjw/Documents/03_Results/Rasters/PH/merged/AL_Y2020_RP100.tif") # for year 2020
AL_rast_50 <- raster("/Users/jeancjw/Documents/03_Results/Rasters/PH/merged/AL_Y2050_RP100.tif") # year 2050

```

```{r}
# Get pixels where there are buildings
AL_rast_20[is.na(building_rast)] <- NA
AL_rast_50[is.na(building_rast)] <- NA
labour_income_rast[is.na(building_rast)] <- NA


# Convert rasters to points 
AL20_pts <- rasterToPoints(AL_rast_20)
AL50_pts <- rasterToPoints(AL_rast_50)
income_pts <- rasterToPoints(labour_income_rast)
building_pts <- rasterToPoints(building_rast)

# Check if 'x' and 'y' coordinates are the same for every row
coordinates_match <- all.equal(AL50_pts[, c("x", "y")], AL20_pts[, c("x", "y")]) # check building_pts as well. This code cannot compare 3 data frames.

if (is.logical(coordinates_match) && all(coordinates_match)) {
  print("Coordinates are the same for every row.")
} else {
  stop("Coordinates are not the same for every row. Check your data.")
}
```

The coordinates for each row are all the same for Year 2050 but not 2020. For 2020, we need to use merge to combine the columns together, but for 2050, cbind will do. 

```{r}
# 2020 asset losses
data_AL20 <- merge(AL20_pts, income_pts, by = c("x", "y")) # merge asset loss with income

data_AL20 <- merge(data_AL20, building_pts, by = c("x", "y")) # merge with buildings

data_AL20 <- data_AL20 %>%
  rename(asset_loss = layer.x,
         income = layer.y, 
         num_buildings = layer)

data_AL20
```
```{r}
# 2050 asset losses
data_AL50 <- merge(AL50_pts, income_pts, by = c("x", "y")) # merge asset loss with income

data_AL50 <- merge(data_AL50, building_pts, by = c("x", "y")) # merge with buildings

data_AL50 <- data_AL50 %>%
  rename(asset_loss = layer.x,
         income = layer.y, 
         num_buildings = layer)

data_AL50
```

Now we want to expand the data frame such that each row represents one building.


```{r}
data_AL20 <- data_AL20 %>%
  select(-x, -y) %>% 
  slice(rep(row_number(), num_buildings)) %>% # replicate rows based on number of households
  select(-num_buildings) 

data_AL20
```
```{r}
data_AL50 <- data_AL50 %>%
  select(-x, -y) %>% 
  slice(rep(row_number(), num_buildings)) %>% # replicate rows based on number of households
  select(-num_buildings) 

data_AL50 
```
```{r}
data_AL20 %>
```

Now we compute the inequity in asset losses expereienced by households of various incomes in the Philippines, under a 1-in-100 year coastal flood in 2020.

```{r}
data_AL20 %>%  
  filter(asset_loss > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=asset_loss, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Asset losses ('000 PHP)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw()
```

We do the same for 2050.
```{r}
data_AL50 %>%  
  filter(asset_loss > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=asset_loss, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Asset losses ('000 PHP)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw()
```

## 5. PLOT WELFARE LOSSES
```{r}
# LOAD DATA
WL_rast_20 <- raster("/Users/jeancjw/Documents/03_Results/Rasters/PH/merged/WL_Y2020_RP100.tif") # for year 2020
WL_rast_50 <- raster("/Users/jeancjw/Documents/03_Results/Rasters/PH/merged/WL_Y2050_RP100.tif") # year 2050

```

```{r}
# Get pixels where there are buildings
WL_rast_20[is.na(building_rast)] <- NA
WL_rast_50[is.na(building_rast)] <- NA
labour_income_rast[is.na(building_rast)] <- NA


# Convert rasters to points 
WL20_pts <- rasterToPoints(WL_rast_20)
WL50_pts <- rasterToPoints(WL_rast_50)
income_pts <- rasterToPoints(labour_income_rast)
building_pts <- rasterToPoints(building_rast)

# Check if 'x' and 'y' coordinates are the same for every row
coordinates_match <- all.equal(WL50_pts[, c("x", "y")], WL20_pts[, c("x", "y")]) # check building_pts as well. This code cannot compare 3 data frames.

if (is.logical(coordinates_match) && all(coordinates_match)) {
  print("Coordinates are the same for every row.")
} else {
  stop("Coordinates are not the same for every row. Check your data.")
}
```

The coordinates for each row are all the same for Year 2050 but not 2020. For 2020, we need to use merge to combine the columns together, but for 2050, cbind will do. 

```{r}
# 2020 asset losses
data_WL20 <- merge(WL20_pts, income_pts, by = c("x", "y")) # merge asset loss with income

data_WL20 <- merge(data_WL20, building_pts, by = c("x", "y")) # merge with buildings

data_WL20 <- data_WL20 %>%
  rename(asset_loss = layer.x,
         income = layer.y, 
         num_buildings = layer)

data_WL20
```

```{r}
# 2050 asset losses
data_WL50 <- merge(WL50_pts, income_pts, by = c("x", "y")) # merge asset loss with income

data_WL50 <- merge(data_WL50, building_pts, by = c("x", "y")) # merge with buildings

data_WL50 <- data_WL50 %>%
  rename(asset_loss = layer.x,
         income = layer.y, 
         num_buildings = layer)

data_WL50
```

Now we want to expand the data frame such that each row represents one building.


```{r}
data_WL20 <- data_WL20 %>%
  select(-x, -y) %>% 
  slice(rep(row_number(), num_buildings)) %>% # replicate rows based on number of households
  select(-num_buildings) 

data_WL20
```

```{r}
data_WL50 <- data_WL50 %>%
  select(-x, -y) %>% 
  slice(rep(row_number(), num_buildings)) %>% # replicate rows based on number of households
  select(-num_buildings) 

data_WL50 
```

Now we compute the inequity in asset losses expereienced by households of various incomes in the Philippines, under a 1-in-100 year coastal flood in 2020.

```{r}
data_WL20 %>%  
  filter(asset_loss > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=asset_loss, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Welfare losses ('000 PHP)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw()
```

We do the same for 2050.
```{r}
data_WL50 %>%  
  filter(asset_loss > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
   ggplot(aes(x=asset_loss, linetype = income_category))+ 
  # stat_ecdf(geom = "step", size = 0.5) +
  geom_smooth(stat = "ecdf", method = "loess", color = "black", se = FALSE, size = 0.5) +  # Apply loess smoothing
  labs(x = "Welfare losses ('000 PHP)", 
       y = "Cumulative frequency", 
       linetype = "Income category") + 
  # scale_color_manual(values = c("low income" = "#cf5529", "high income" = "#294475"))+
   scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  theme_bw()
```

## 6. Combine ASSET & WELFARE loss

```{r}
# categorise data
data_AL20 <- data_AL20 %>% 
  mutate(loss_category = "Asset loss", 
         year = "2020")%>% 
  rename(losses = asset_loss)
data_AL50 <- data_AL50 %>% 
  mutate(loss_category = "Asset loss", 
         year = "2050")%>% 
  rename(losses = asset_loss)
data_WL20 <- data_WL20 %>% 
  mutate(loss_category = "Welfare loss", 
         year = "2020")%>% 
  rename(losses = asset_loss)

data_WL50 <- data_WL50 %>% 
  mutate(loss_category = "Welfare loss", 
         year = "2050") %>% 
  rename(losses = asset_loss)

```

```{r}
loss_data_20 <- rbind(data_AL20, data_WL20)
loss_data_20 <- bind_rows(mutate(loss_data_20 %>% 
                   filter(loss_category == "Asset loss"),
                 loss_category = factor(loss_category), 
                 colour = "#cf5519"),
          mutate(loss_data_20 %>% 
                   filter(loss_category == "Welfare loss"), 
                 loss_category = factor(loss_category),
                 colour = "#294475"))

```

```{r}
loss_data_50 <- rbind(data_AL50, data_WL50)
loss_data_50 <- bind_rows(mutate(loss_data_50 %>% 
                   filter(loss_category == "Asset loss"),
                 loss_category = factor(loss_category), 
                 colour = "#cf5519"),
          mutate(loss_data_50 %>% 
                   filter(loss_category == "Welfare loss"), 
                 loss_category = factor(loss_category),
                 colour = "#294475"))

```

### 6.1 Year 2020
```{r}
loss_data_20 %>%
  filter(losses > 0) %>%
  ggplot(aes(x = losses, color = loss_category)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Loss type") +
  # scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```

Asset vs welfare losses, not differentiated by income. 

```{r}
loss_data_20 %>%
  filter(losses > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  ggplot(aes(x = losses, linetype = income_category, color = loss_category)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Loss type") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```
Here, we observe that for those below the poverty line, asset losses are very few. 

```{r}
loss_data_50 %>% 
  filter(losses > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Below poverty line") %>% 
  ggplot(aes(x = losses, linetype = income_category, color = loss_category)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Loss type") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```

### 6.2 Year 2050

```{r}
loss_data_50 %>%
  filter(losses > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  ggplot(aes(x = losses, linetype = income_category, color = loss_category)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Loss type") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```

```{r}
loss_data_50 %>% 
  filter(losses > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Below poverty line") %>% 
  ggplot(aes(x = losses, linetype = income_category, color = loss_category)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Loss type") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```
```{r}
loss_data_50 %>% 
  filter(losses > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Above poverty line") %>% 
  ggplot(aes(x = losses, linetype = income_category, color = loss_category)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Loss type") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```
## 7. Effect of SLR
### 7.1 Asset losses

Compare AL 2020 and 2050; WL 2020 and 2050

```{r}
asset_loss_data <- rbind(data_AL50, data_AL20)
asset_loss_data  <- bind_rows(mutate(asset_loss_data %>% 
                   filter(year == "2020"),
                 year = factor(year), 
                 colour = "#cf5519"),
          mutate(asset_loss_data %>% 
                   filter(year == "2050"),
                  year = factor(year), 
                 colour = "#294475"))

asset_loss_data
```
```{r}
welfare_loss_data <- rbind(data_WL50, data_WL20)
welfare_loss_data  <- bind_rows(mutate(welfare_loss_data %>% 
                   filter(year == "2020"),
                 year = factor(year), 
                 colour = "#cf5519"),
          mutate(welfare_loss_data %>% 
                   filter(year == "2050"),
                  year = factor(year), 
                 colour = "#294475"))

welfare_loss_data
```


```{r}
asset_loss_data %>%
  filter(losses > 0) %>%
    # mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  ggplot(aes(x = losses, 
             # linetype = income_category, 
             color = year)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       # linetype = "Income category",
       color = "Year") +
  # scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```

```{r}
asset_loss_data %>%
  filter(losses > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>%
  ggplot(aes(x = losses, 
             linetype = income_category,
             color = year)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Year") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```
### 6.2 COMPARE WELFARE LOSSES (BY YEAR)

Welfare losses are generally more pronounced in 2050 than in 2020.

```{r}
welfare_loss_data %>%
  filter(losses > 0) %>%
    # mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  ggplot(aes(x = losses, 
             # linetype = income_category, 
             color = year)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       # linetype = "Income category",
       color = "Year") +
  # scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```



```{r}
welfare_loss_data %>%
  filter(losses > 0) %>%
    mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>%
  ggplot(aes(x = losses, 
             linetype = income_category,
             color = year)) +
  geom_smooth(stat = "ecdf", method = "loess", se = F, size = 0.5) +
  labs(x = "Losses ('000 PHP)",
       y = "Cumulative frequency",
       linetype = "Income category",
       color = "Year") +
  scale_linetype_manual(values = c("Below poverty line" = "dashed", "Above poverty line" = "solid")) +
  scale_color_manual(values = c("#cf5519", "#294475")) +
  theme_bw()
```
However, when we view this based on income groups, we can see that those living below the international poverty threshold will experience a larger difference in welfare losses due to sea level rise. 

We confirm these findings by computing the maximum vertical distance between the solid lines, and the max vertical distance between the dotted lines. 

#### 6.1.3 Maximum vertical distance



```{r}
welfare_loss_data_low %>% 
  filter(losses > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Below poverty line") %>% 
  arrange(depth)
high_income20 <- data20 %>% 
  filter(depth > 0) %>% 
  mutate(income_category = ifelse(income <= pov_line_intl, "Below poverty line", "Above poverty line")) %>% 
  filter(income_category == "Above poverty line") %>% 
  arrange(depth)


cdf_low <- ecdf(low_income20$depth) # compute cdf values for the households below poverty line
cdf_high <- ecdf(high_income20$depth) # compute cdf values for the households above poverty line

depth_values <- seq(min(data$depth), max(data$depth), length.out = 1000)
differences <- cdf_high(depth_values) - cdf_low(depth_values)
abs_difference <- abs(differences)
l1norm50 <- sum(abs_difference)
cat("L1 Norm:", l1norm50, "\n")


max_difference <- max(differences)
max_difference_depth <- depth_values[which.max(differences)]


cat("Maximum Difference:", max_difference, "\n")
cat("Corresponding Flood Depth:", max_difference_depth, "\n")
```

